### 2020-07-01 writing/experiment.tex
paper, the first blood

\section{The Experimental Setup}

\begin{itemize}

\item Summary of the blame trail experiment. We sample and follow blame trails through configuration
        lattices.

\item Evaluating each point in the design space. We run the blame trail experiment for every point in the
        previously discussed design space.

\item Selection of programs. We are using a set of representative
  programs. Argue the programs are representative of the range of programs written by TR developers; point to the JFP paper.

\item Sampling of mutants. We have too many mutants to run them all, so we sample them with an aim to represent equally all of the mistakes captured by our operators.

\end{itemize}