### 2020-07-01 writing/ideas.tex

\section{How to Evaluate Blame}

\begin{itemize}

    \item How to analyze this space?
          \begin{itemize}
            \item We want a quantative measure of error msg utility that allows
                  us to evaluate each point in the above space.
          \end{itemize}

     \item Blame trails
       \begin{itemize}

        \item We want an empirical study to understand how the behavior of blame
        in real systems.

        \item Blame trails provide a quantitative measure of a a system's
        error message utility, and they can be applied to every point
        in the above space.

       \item We generalize the blame trails idea of Lukas et al [POPL '20] to make it
        more systematic, scalable, and general, by:
           \begin{itemize}
             \item Adding a null hypothesis baseline for comparison between multiple systems
             \item Analyzing multiple blame following strategies
             \item Mutating interfaces as well as programs
             \item Incorporating domain knowledge in the selection of blame following
           strategies and mutation
             \item Evaluating systems on the same platform for apples-to-apples comparisons
           \end{itemize}

         \item Null hypothesis. We should have a baseline to tell how all of the points in the
        design space compare to having nothing at all.
\end{itemize}