### 2020-07-01 writing/paper.tex

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[acmsmall,screen]{acmart}\settopmatter{printfolios=true,printccs=true,printacmref=true}
%% For double-blind review submission, w/ CCS and ACM Reference
%%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from PACMPL format to traditional
%% SIGPLAN proceedings format must update the '\documentclass' and
%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Title information
\newcommand{\thetitle}{How to Evaluate Blame for Gradual Types}
\title{\thetitle}

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%\author[]{}
%\affiliation{
%\institution{}
%% \city{}
%% \city{}
% \country{}
%}
%\email{}          %% \email is recommended

\usepackage{alltt}
\usepackage{wrapfig}

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\usepackage{listings}
\usepackage[shortcuts]{extdash}
\usepackage{xcolor}

\usepackage{upgreek}

\input{def.tex}

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
%% \setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

%%%
\setcopyright{}
\acmPrice{}
\acmDOI{}
\acmYear{}
\copyrightyear{}
\acmJournal{}
\acmVolume{}
\acmNumber{}
\acmArticle{}
\acmMonth{}

\begin{document}

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
Gradual types and blame go hand in hand. But what is the value of blame?
\end{abstract}

% CCS concepts: Program specifications < Program reasoning < Semantics and Reasoning < Theory of Computation
% CCS concepts: Empirical software validation < Software verification and validation < Software creation and management < Software and its engineering

% potential choices decided against: pre- and post-conditions, assertions, software testing and debugging

%\begin{CCSXML}
%<ccs2012>
%<concept>
%<concept_id></concept_id>
%<concept_desc></concept_desc>
%<concept_significance></concept_significance>
%</concept>
%<concept>
%<concept_id></concept_id>
%<concept_desc></concept_desc>
%<concept_significance></concept_significance>
%</concept>
%</ccs2012>
%\end{CCSXML}

%\ccsdesc[]{}
%\ccsdesc[]{}

%% Keywords
%% comma separated list
%\keywords{}  %% \keywords are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\input{introduction.tex} \label{sec:introduction}
\input{landscape.tex} \label{sec:landscape}
\input{ideas.tex} \label{sec:ideas}
\input{experiment.tex} \label{sec:experiment}
\input{results.tex} \label{sec:results}
\input{discussion.tex} \label{sec:discussion}
\input{related.tex} \label{sec:related}
\input{conclusion.tex} \label{sec:conclusion}

%% Acknowledgments
%%\begin{acks}                 
  %% acks environment is optional
  %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
%%\end{acks}

%% Bibliography
\bibliography{cs}

\end{document}