### 2020-11-11 writing/paper.tex
format + prose tweaks

\documentclass[sigplan,10pt,review,anonymous]{acmart}
\settopmatter{printfolios=true,printccs=false,printacmref=false}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from PACMPL format to traditional
%% SIGPLAN proceedings format must update the '\documentclass' and
%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Title information
\newcommand{\thetitle}{How to Evaluate Blame for Gradual Types}
\title{\thetitle}

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

\author{Nobody In Particular}
% \affiliation{\institution{PLT} \city{Evanston} \country{Illinois} }
\affiliation{\relax} % \institution{PLT} \city{Evanston} \country{Illinois} }
% \email{chrdim@northwestern.edu}          %% \email is recommended

\usepackage{alltt}
\usepackage{wrapfig}

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\usepackage{listings}
\usepackage[shortcuts]{extdash}
\usepackage{xcolor}

\usepackage{upgreek}

\input{def.tex}

\acmConference[]{}{}{}
\acmYear{}
\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}

\begin{document}

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
Programming language theoreticians develop blame assignment systems
and prove blame theorems for gradually typed programming languages.
Practical implementations of gradual typing almost completely ignore
the idea of blame assignment.  This contrast raises the question
whether blame assignment provides any value to the working programmer
and poses the challenge of how to evaluate the effectiveness of blame
assignment strategies. This paper contributes (1) the first evaluation
method for blame assignment strategies and (2) the results from applying
it to three different gradual type systems. \mf{This sentence must be checked
before final submission.} These results confirm that blame assignment
is helpful to a rationally acting programmer.
\end{abstract}

% CCS concepts: Program specifications < Program reasoning < Semantics and Reasoning < Theory of Computation
% CCS concepts: Empirical software validation < Software verification and validation < Software creation and management < Software and its engineering

% potential choices decided against: pre- and post-conditions, assertions, software testing and debugging

%\begin{CCSXML}
%<ccs2012>
%<concept>
%<concept_id></concept_id>
%<concept_desc></concept_desc>
%<concept_significance></concept_significance>
%</concept>
%<concept>
%<concept_id></concept_id>
%<concept_desc></concept_desc>
%<concept_significance></concept_significance>
%</concept>
%</ccs2012>
%\end{CCSXML}

%\ccsdesc[]{}
%\ccsdesc[]{}

%% Keywords
%% comma separated list
%\keywords{}  %% \keywords are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\input{introduction.tex} 
\input{challenges.tex}   
\input{landscape.tex} 
\input{mutate.tex}
\input{rational.tex}	 \label{sec:rational}
\input{sample.tex}
\input{results.tex}      \label{sec:results}
\input{discussion.tex}   \label{sec:discussion}
\input{related.tex}      \label{sec:related}
\input{conclusion.tex}   \label{sec:conclusion}

%% Acknowledgments
%%\begin{acks}                 
  %% acks environment is optional
  %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
%%\end{acks}

%% Bibliography
\bibliography{cs}

\end{document}