### 2021-02-11 writing/reviews/pldi-rebuttal-final.md
Combine top and bot into -final.md

PLDI 2021 Paper #66 Author Response
===========================================================================

Thank you to all the reviewers for the helpful feedback.

The first part of our response addresses two common themes in the
reviews; the second part addresses all remaining concerns individually.

Are the empirical results conclusive? Are they surprising?
===========================================================================

Here are the two relevant major conclusions from our experiment:

1. Transient and Natural's blame makes rational programmers more effective than Erasure.

2. The effectiveness of Natural is (merely) comparable to Transient.

As review `D` points out, the second one is surprising: _While Theory
predicts that Natural blames more accurately than Transient, our
experimental results do not support this prediction._

In analogy to physics research, this points to two possible explanations:

- a need for theoretical models with improved predictive power

In all likelihood, the current models are too small to include language features
critical to blame assignment.

- a need for "blame data" from large systems

The presented data is based on systems with up to 14 modules (see table
below).  Natural may prove its worth only when the dependency chains are
deeper than the ones in the existing benchmarks. If Natural works better
for programs with deeper dependency chains, this should guide future research
on mixing approaches.

Is the method novel? 
===========================================================================

Our inspiration is Lazarek et al.'s POPL paper, which shows the possibility
of large-scale design evaluations that do _not_ involve human software devs.

The challenge is how to adapt this work to problems other than
contracts. Given the close relationship between Natural and contracts, it
is "natural" to pursue an adaptation of the Lazarek method to gradual
typing. Adaptation means retaining some ideas and injecting entirely new
ones:

1. Lazarek et al.'s work can use off-the-shelf mutators because they are only
   looking for bugs that contracts can discover in an untyped world.  Our
   work demands an entirely new judgment of which mutators are "good" or
   "bad". This new judgment means cutting one mutator, changing four of them,
   and adding nine. There was no other way to get an _interesting_ and
   diverse body of bugs (see section 4).

2. Lazarek et al. deal with a single contract system. Our experiment compares
   fundamentally different checking and error-reporting schemata: the
   industrial Erasure standard and the academic ones (Natural and Transient).

3. Lazarek et al. examine a single yes-or-no question: whether blame-shifting
   works. Our work asks (a) several yes-or-no questions and (b) asks
   questions of degree; it yields statistically sound answers for the former
   and evidence to shed light on the latter.

4. Lazarek et al.'s method cannot really conclude that blame is the reason
   for the success of the debugging process, because it fails to isolate the
   effect of blame from that of checks or mere luck. It simply validates that
   blame shifting works. Our work focuses on the value of blame and that kind
   of work demands proper controls, which are entirely new.

   Our work relies on two different baselines to understand the value of blame:

   - _the "random" ("lottery") programmer_ ~~ An irrational programmer ignores
     the error report and adds types to one randomly chosen component at a
     time. Hence we use the "random mode" as a baseline to show that following
     the error report is _not_ a question of luck.

   - _the run-time checking of the underlying language_ ~~ All sound gradual
     type systems employ run-time checks to guarantee the consistency of types,
     but they do not have to track blame. If a run-time check fails, they can
     raise an exception that reports a stack (trace).  Many programmers use
     stack traces for debugging, but our work must show that blame assignment
     is better than keeping checks and throwing out blame.

5. One additional point from review `D` stands out, namely, the remark that
   bringing Natural and Transient under the same roof (section 3) is "just
   engineering."

   Neither Reticulated Python nor the model of Transient provide answers
   about what shallow checks correspond to the sophisticated types of Typed
   Racket (or TypeScript), e.g.  polymorphism, row types for mixins,
   occurrence typing) and what information is necessary to record in the
   blame map.  But, we do not think that this paper is the proper place to
   describe this aspect of the work; the large explanation would distract the
   reader from the actual thesis of the paper.

Replies to individual points
==================================================================

Review #66A
===========================================================================
> - Could you comment a bit more on what makes a mutator "good"? (other
>   than experimentally seeing that it is good.) Perhaps, what would be a
>   really bad mutator for evaluating blame assignment?

A bad mutator results in a type-level bug that is trivial to locate. For
instance, a bug that leads to a type-level error during the evaluation of the
component that contains it is trivial to locate. In contrast, a bug that is
non-trivial to locate is one that causes the buggy component to evaluate without
an error and to produce a value of the wrong type. An interesting bug is
one where the value with the wrong type manages also to cross undetected to at
least one other component before a type-level check flags it. Good mutators
should be able to produce these interesting bugs.

> - I don't understand where the 16,800 number comes from. Is it an
>   exhaustive enumeration of all possible mutations? (Or the number of
>   mutants constructed in X amount of time?)

It is an exhaustive enumeration. We will clarify the prose.

> - How large are the programs in the benchmark suite?

Here are the LOC counts and number of components for each benchmark. We
will add the information to figure 2.

  benchmark    | LOC  | components
  -------------|------|-----------
  acquire      | 1941 | 9
  gregor       | 2336 | 13
  kcfa         | 328  | 7
  quadT        | 7396 | 14
  quadU        | 7282 | 14
  snake        | 182  | 8
  suffixtree   | 1500 | 6
  synth        | 871  | 10
  take5        | 465  | 8

> # Minor Comments
> - L38: Should there have been a colon? Is the text "This contrast... "
>   and challenge quote from [3]? Who is saying what here...

No, that text is not a quote from [3]. We will rephrase to clarify that it is
our statement of the research question.

> - L104: I was confused by the word "Next". Maybe instead: We now ..."

Thank you. We will reword.

> - L268: Grammar issue.
> - L314: Floating period.
> - L441: Add line numbers (or other metric of size).
> - L980: Figure 5.4??

We will fix these. Thanks.

> - L956: 30,000 hours of compute time is 1,250 days. I assume that the
>   CPU has many cores. Could you put that number?

Each CPU has 28 doubly-threaded cores. We will add it to the line.

> - L991: Figure 6: It took me some time to understand the figure. I
>   wonder if there is a better way to represent the same data?

We have reached the same conclusion and we are considering alternatives
that will make it easier to see how two modes compare to each other, such
as a series of two-sided bar charts.

Review #66B
===========================================================================
> The experiment results show that there are debugging scenarios where
> Transient is more useful than Natural. Does it happen simply because of
> luck? Or, does it imply that Transient is more suitable than Natural for
> a particular kind of bugs for some reason? Also, can there be a new
> blame strategy that outperforms both Transient and Natural all the time?

We were also surprised to discover that Transient outperforms Natural in
some scenarios. After all, theory says otherwise. We haven't identified
any patterns for these scenarios. We conjecture that they point to
limitations of the theory but we can't exclude that some of them aren't
because of some unknown Typed Racket bug.

We do not have any leads for a new blame strategy better than both
Transient and Natural, but at least our method provides a way to examine
ideas in that direction.

> Theoretical gradual type systems usually allow programmers to control
> the precision of type annotations in a fine-grained way with the Dyn
> type rather than distinguishing typed and untyped modules. Would the
> same evaluation method be able to directly appliable to blame in such
> systems?

Yes. Incorporating type Dyn boils down to considering a hierarchy of types
for each component rather than a single type. At the same time, our method
is parametric to the "size" of a component and adapts in a straightforward
manner to a setting where each definition can be considered a component
instead of a whole module. Of course both of these adjustments lead to
significantly larger scenario lattices than the ones we investigate, so
keeping the experiment computationally feasible would require more
aggressive sampling.

> Please explain what impedance mismatches are in the paper. Is it simply
> type mismatches?

"Impedance mismatch" is our name for a run-time disagreement between a
type annotation and an untyped value. The point of the name is to suggest
that either side (type or value) could be at fault. We will adjust the prose.

> The paper states that the effort distribution of the random mode follows
> a normal distribution, as expected. Please explain why a normal
> distribution is a reasonable expectation. ...  This point does not
> affect the contribution of the paper, but it is not clear from the
> paper.

You are right that for a given trail of length $n$ the probability that
the random rational programmer discovers the buggy component on every step
is $1/n$. The expected curve is not flat though, because we randomly
sample from the set of scenarios --- and by extension the set of trails.
That said, as you observe, the distribution of the random programmer's
effort is not relevant to our results, so we will remove the
characterization.

> * page 9 line 980: Figure 5.4 -> Figure 6
> * page 10 Figure 7: Null -> Random
> * page 11 line 1103: ( 770 -> (770

Thank you. We will fix all these.

Review #66C
===========================================================================
> But while there is a lot to like about the paper, it also has some
> shortcomings. The suite of programs used for the evaluation is fairly
> small (c.f., the SIGPLAN checklist).

The paper does follow the SIGPLAN checklist guidelines for Principled Benchmark
Choice:

- This paper is the first to conduct an empirical study for blame and gradual
  typing, so there is no established benchmark suite to reuse.

- The paper thus explains how we adapted a suite of gradual typing performance
  benchmarks to a new setting (section 4).

- The paper also explains how we selected pertinent benchmarks to work within
  our resource constraints.

We cannot find any minimum size guidelines in the SIGPLAN checklist.

> The faults are introduced using synthetic mutations that may or may not
> correspond to the kinds of errors that arise in practice. This can be
> seen from the results in Figure 7 -- the lengths of the "trails" is
> quite small.

There is no existing catalog of bugs in gradually typed programs, nor any
statistics about common trail lengths.  Furthermore, most type-level bugs do
not survive  decent testing and thus do not make it to code repos. There are a
few anecdotes from the literature but they are mostly artificial. Hence, since
we need a large number of buggy scenarios for a meaningful study, a synthetic
approach is the only way forward. In addition, mutation allows us to tune the
set of scenarios to get a diverse set of interesting bugs.

> While the study is mostly well done, it only considers two blame
> assignment strategies.

We would like to point out that the experiment considers six different
modes of the rational programmer spread across three different gradual typing
systems plus the random rational programmer. Figure 7 clearly shows them all.

> For example, it would be interesting to consider the "omniscient
> programmer" to get a bound on the value of blame. But this is not done.

If an "omniscient programmer" is the programmer that always makes the best
choice, i.e. finds the bug at the first try, then this corresponds to a
trivial mode; it has 100% success rate and a constant blame trail length
of 1. Thus, comparing its success to any other strategy amounts to
counting the number of scenarios where the other strategy fails; comparing
effort is similarly trivial. This information can be inferred from the
definitions of section 5, but we can certainly add an explicit mention if
it would be useful.

> And unfortunately the results are inconclusive -- Section 8 says "our
> results call for a deeper understanding of the two models for blame".

The quote from section 8 refers to the theory of blame rather than our
results. Please see the first part of the discussion at the beginning of
the response for more details on the conclusions of our results.

That said, the results of the comparison between the three systems in
section 7 is only one part of the contribution of the paper. The other
part is the method itself (sections 3, 4, and 5).

> Also, the published Transient scheme seems to have a minor bug, which
> might affect the outcomes when fixed (Section 8.2).

Section 8.2 does not discuss a bug in Transient but rather a limitation of
its current design that we discovered exactly because of the experiment.

> To explain one comment above: I was surprised not to see a comparison
> against an "omniscient programmer" to establish a baseline (c.f., the
> SIGPLAN checklist). That is, model a blame assignment scheme in which
> blame could be placed on any component. And always pick the component
> that minimizes the length of the trail. Of course, this would be
> infeasible to implement in practice, but it would place an upper bound
> on the utility of blame.

Our experiment has several baselines. The random mode of the rational
programmer described in section 5.5 serves as the "control" of the whole
experiment. In addition, the exception modes described in sections 5.1 and
5.2 each serve as further "controls" to distinguish the effect of checks
from that of blame within each semantics. Please also see our discussion
of baselines in the second theme at the beginning of our reply.

> Similarly, I wondered why the modes of the transient programmer only
> consider picking the first/last blame, and not other choices. Section
> 5.2 just says "our answer" is that there are at least "two reasonable
> options." Why are other choices not reasonable?

The Transient semantics provide no interpretation for the blame set. We
picked the first and last because they match the intuitive interpretation
of the blame set as a list. We didn't mean to imply that other options are
unreasonable. We just needed to make a reasonable choice to keep the
experiment feasible.

Review #66D
===========================================================================

> 2. That Transient --- as implemented in Typed Racket --- is much
> slower than Natural. ...
> Given that the experiments are developed and run in the context of the
> large, real-world Racket ecosystem, I wonder whether about the chance of
> incidental implementation choices or bugs affecting this result.

A side note here: Transient Typed Racket is faster than Natural Typed
Racket when we turn off blame. As a sanity check that this is not an
artifact of our implementation, we translated some benchmarks to Python
and confirmed that in Reticulated the performance of Transient with blame
is even worse than we have observed in Typed Racket.

> *Custom Mutators (Section 4)* The paper suggests that the mutators
> required here are significantly different and more subtle, but this does
> not seem to be the case when comparing Figure 3 and Table 3 of the POPL
> 2020 paper.  ...

First of all, we need to clarify that the mutators used by Lazarek et al.
are standard mutators from the the extensive literature on mutation
testing. Naturally, we started with those to develop our mutators, but we
quickly realized that only three of them work for our purposes. In
hindsight, that is not surprising because they are designed to cause
behavioral bugs rather than type-level ones. This realization led us to
think carefully about interesting bugs in our setting and develop a new
set of mutators based on Lazarek et al.'s that meets our needs. For
example, the relational mutator from Lazarek et al. does not lead to
type-level errors that Typed Racket's type system can detect, so we
eliminated it. Similarly, the original arithmetic mutator does not
reliably lead to type errors, so we adapted it to only make operator swaps
that can affect the type of the result. For example, changing a `/` to `*`
will not affect the type of an expression, but changing `*` to `/` may! In
addition to those adjustments, we had to come up with entirely new
mutators in order to get a sufficiently large and diverse set of bugs. For
example, the new `-id` family of mutators are perfectly suited to creating
type-level mistakes, and in fact most of our interesting bugs are
generated by them. These mutators are completely different than swapping
subexpressions; in particular, they are context-sensitive and nonlocal.
Finally, we found mutators that we expected should do well but turned out
to be ineffective at producing interesting bugs. For example, adding a new
public method to a class reliably produces both static and dynamic type
errors, but we found they were not interesting bugs.

Besides all of these details about the mutators, we consider the biggest
insight that allowed us to find the right set of mutators to be the
definition of interesting bugs. Please see the high level comments at the
start for more discussion of this.

> *Debugging Strategies (Section 5)*
>
> These also seem to be overemphasized. The main design choice seems to be
> in how to use Transient blame, which reports multiple components, and
> picking the first and last are pretty intuitive.
>
> I'm also not sure why the exception mode is needed in addition to
> Erasure (I was confused in Section 2 L204-215 and Section 5.1 L732).
> Indeed, Section 5.3 defines the Erasure mode the follow the Natural
> exceptional mode.

The key point of section 5 is the definition of the different blame
strategies in a common precise framework as modes of the rational
programmer. Modes enable the comparison between the different systems even
when sampling is involved, thanks to the common substrate of blame trails.

In addition, modes allow us to introduce in a uniform manner in the
framework baselines that help us isolate confounding factors from the
actual effectiveness of blame. This is exactly the role of the exception
modes for Natural and Transient. Those two perform checks at runtime and
the exception modes help us determine what part of the success of Natural
and Transient is due to the extra checks or due to blame (and also whether
blame masks useful information from exceptions). For example, we see that
Natural without blame but with the same checks outperforms Erasure in ~25%
of the scenarios (figure 6). Blame adds another ~11% on top of that. For
Transient there is a small percentage of scenarios where exceptions do
better than blame but still blame improves over exceptions by ~11%
compared to Erasure.

> *Additional Comments and Typos*
>
> L69: I wish here, and elsewhere, the comparison to [13] had been made
> more explicit. As discussed above, there seem to be more similarities
> than suggested.
> L92: "use of _a_ higher-order contract system"
> L155: "authors extensive"
> L253: "exports _it_ as"
> L280: Missing `untyped-`
> L314: " . "
> L382: "Dyn,"
> Fig 2: Presumably authors were included to indicate the variety of
> sources, but I'm not sure that was necessary.
> L597: "make us_e_ of"
> L664: "described in _Section_ 4"
> L980: "Figure 5.4" ==> "Figure 6"
> L986: "2.8%" ==> "2.18%" ?
> L1017: "33%" ==> "26%" ?
> L1304: "Problem is ...."

Thank you. We will fix these.

Review #66E
===========================================================================

> However, I don't find the results to be as conclusive as the authors
> state. Aside from the Erasure case, whose relevance I found unclear (see
> below), conclusions rest on some small-valued "more useful than"
> percentages. These are pretty hard to interpret, because the
> "usefulness" metrics necessarily build in a lot of simplifying
> assumptions. E.g. in 5.4, the "percentage of scenarios where ... is more
> effective" doesn't account for *extent* of difference, and the
> trail-length "programmer effort" metric in 5.5 is also unlikely to be a
> great proxy. These limitations are understandable, method-wise, but
> overall the results are fairly null... they let us continue believing
> "what we'd expect" but not with any great sense of added confidence.
> That doesn't entirely diminish the work, of course, but I'd say the
> current write-up overplays its hand a little. I appreciate the
> discussion of threats to validity.

Please see the discussion at the beginning of our response about the
interpretation of our results, their significance, and how they challenge
existing perceptions about blame.

We are not sure what "extent" means in this comment. Usefulness asks
whether one system succeeds in a debugging scenario while another system
fails. If both succeed we use the length of the trail to quantify the
difference between the two systems. Trail length may not be the most
optimal proxy for programmer effort, but it also allows us to perform a
large scale automatic experiment about blame without the practical
limitations of user studies. We will add a discussion of this to the
paper.

> The presentation of the results is very much around aggregates and summaries.
> Indeed the whole method is about having run a huge compute job over a large
> number of variants. Perhaps it's paranoia but I'd be interested to see some
> specific examples walked through by hand (in the paper) and some smaller sample
> manually classified (as results). That would add an extra sanity check that the
> metrics do correspond to some meaningful reality.

We are not sure what the question is here. If the question is whether our
implementation of the experiment is faithful to the definitions from
section 5 and the discussion about statistics from section 6, then the
answer is emphatically yes. We have validated manually a number of blame
trails for each mode and moreover our implementation of the experiment
comes with plenty of sanity checks.

If the question is whether our methods acts as a correct classifier for
the detection of bugs a la machine learning, then we would like to stress
that this is not what our work is about. Our work defines the rational
programmer, an ideal programmer, as a deterministic algorithmic procedure
and then uses the results of this procedure on a large number of debugging
scenarios to determine if blame is a critical factor for the success or
failure of the rational programmer.

> About Erasure: if I'm reading this correctly, the key thing here is that
> in Typed Racket, the set of static types is more expressive than the set
> of dynamic types. For example, there is a static notion of non-negative
> integer that is distinct from plain integer. This sort of design isn't
> universal -- in some languages/systems static types closely mirror the
> classification of objects in the language's dynamic semantics. Writing
> these stricter contracts into a program brings its own benefits,
> separate from blame or indeed from static checking. It feels like the
> paper doesn't take enough care to distinguish the two effects: the
> effect of systematically applying more refined contracts over program
> values, and the effect of gradually enabling static checking of those
> contracts (iteratively, guided by blame). The authors do mention this
> around line 204, and the comparisons in Figure 6 between "_ exceptions"
> and "Erasure" seem to be measuring this -- the gain from checking these
> extra annotations, with the more precise and/or more timely checks that
> they imply, relative to the erased case where only the language baseline
> contracts are checked. Indeed that's the point of 'exception'
> experiments. Since the biggest effect sizes on display are these ones --
> between Erasure and anything else -- this seems at best distracting. I'd
> be glad to hear from the authors if I'm misunderstanding anything here.
> In the detailed comments below, I have noted some places in the text
> where it would be useful to remind the reader that this design property
> of Typed Racket is at play.

We think there might be a misunderstanding here about gradual typing:

1. Erasure is the established name of the standard semantics implemented
   by industrial languages such as TypeScript, Flow etc. These languages
   come with a type system just as sophisticated as Typed Racket.
   Programmers add annotations to their program and the type checker does
   its best to type check it in order to discover mistakes and help the
   IDE. Then types are erased and the program is run in the underlying
   dynamically typed language. That language's primitive operations have
   runtime checks that may detect an incompatible argument and raise an
   exception.

2. The two academic semantics introduce extra run time checks (in the form
   of proxies for Natural and inlined checks for Transient, as section 3
   describes). The checks enforce that values the untyped code provides to
   typed code behave according to their expected type. When one of these
   checks fails both Natural and Transient issue blame (though each with
   different mechanisms).

3. In Typed Racket, programmers cannot add contracts. The precision of the
   checks matches the type annotations of the program.

4. The Natural and Transient exception modes perform the same checks as
   Natural and Transient correspondingly. The only difference is that they do
   not produce blame but instead signal an exception with the stack trace of
   the failed check. We use these modes to separate the effectiveness of
   blame from that of the checks. Please also see the second part of the
   discussion at the beginning of our response.

> It's interesting to note that there seems to be no convincing analogous
> experimental or modelling-based justification for plain old static
> typing, i.e. showing that it somehow presents a net gain to the
> programmer. Rather, this has simply been posited/assumed by a very long
> line of work. I am not defending that state of affairs, but it points to
> the difficulty of showing conclusively that something truly helps
> programmers. I can see value in this sort of simulation-style approach,
> and it is no less convincing than user studies. So I'd be interested to
> hear from the author(s) if they have any more arguments that (1) I've
> underestimated the results' conclusiveness, or (2) this is a
> novel/interesting family of methods that might be pursued more widely,
> or (3) that deeper experiments building on these ideas might yield more
> compelling insights.

For points 1 and 3 please see our discussion at the beginning of the
reply. For point 2, the idea of the rational programmer is a new one so
far applied only to blame in contracts (Lazarek et al. POPL 2020) and blame in
gradual typing (our work). We do believe it can have a similar effect in
programming languages as homo economicus in economics.

That said, as posed, this comment seems to express a significant
mis-characterization of the submission. It is not the benefits of a
gradual type system that are in doubt. The experimental setup keeps the
type system constant but allows to answer the question which of several
checking and blaming regimes (for enforcing types) provides the best
explanatory messages in case of violations. Hence the analogous question
for purely static type systems would ask which of several reporting
schemes provides the best explanatory message for type errors. In the case
of simply typed languages and even languages with local type inference,
this question is basically meaningless. In the case of languages with HM
type inference, the question has been implicitly raised for four decades
with the development of alternative ways of finding the source of
inference conflicts. One recent example is SHErrLoc (Zhang et al. TOPLAS
2017).

If the reviewer is indeed interested in the question of whether type systems
help programmers --- a question that this submission does _not_ ask --- then
Hanenberg and Stefik's papers are a good starting point (e.g. ICSE 2014).

> line 34: "then their compilers remove types and rely on the built-in
> safety checks of the underlying language to catch any problems". This
> reads oddly in context. Didn't they just do a bunch of static checking?
> So they are *not* relying just on dynamic checks to catch problems?
> Maybe there is something more accurate to say here... e.g. no dynamic
> check is removed, or something like that.

In Erasure, a program can type check and still have type-level errors
because some type annotations may be missing. Section 3 illustrates
exactly this scenario by example. In contrast to Natural and Transient,
types in Erasure do not result in run time checks so the only way to
detect type-level errors that the type checker could not find is the
built-in checks of the language's primitives.

> line 38: "explicit statement and challenge" -- what is it?

The challenge questions whether blame is useful at all in gradual typing.
We will clarify this in the prose.

> line 49: I agree they got the word wrong, but you should explain this

> line 60: around this paragraph the writing started to grate. There's no
> need to generalise about what people do or think, and the "As a matter
> of fact... simply..." style is somewhere between laboured and
> patronising. It's better to more plainly state the gap in the literature
> that you're addressing. Throughout the paper, much space could be saved
> by writing in a more direct style.

Noted, we will rephrase.

> line 69: be explicit that Lazarek et al were (as I later gathered) doing
> something about blame to do with higher-order contracts but not gradual
> types

Thanks, we will add a more detailed comparison along the lines of the
discussion at the beginning of our response.

> line 70: what does it mean to "follow" the slogan? Was unclear to me.

The sentences that follow line 70 are meant to expand on the slogan. We
will adjust the prose to make it clear.

> line 90: at first I wondered: what is a case? Maybe say "program
> variant" to foreshadow the idea of generating mutants etc?

It means debugging scenario. We will clarify the prose.

> line 91-ish: "Transient, "Natural", "Erasure" -- be explicit that these
> are names that *you* are introducing

We are reusing these names from Greenman et al. and Vitousek et al. [10, 37],
but will explain more.

> line 187: what is a "component"? Should be easy to define.

In the context of Typed Racket, a component is a module.

> line 277: what is "responsibility" of a "party", exactly? Are "party",
> "module" and "component" all the same things?

Yes in this context. Party and component refer to the entities that can be
blamed, which in Typed Racket is modules.

> line 293: I was wondering what constitutes a boundary crossing. Clearly,
> passing by function call or return crosses from the caller's module to
> the callee's. What about values exchanged through reads/writes to shared
> state?

Yes, exchange of values via state also constitutes a boundary crossing.
Some of our benchmarks do exactly that. We will clarify what a boundary
crossing is in the prose.

> line 313: it threw me that a program might not fail but "produce a wrong
> result". If it could be caught by the gradual type system, why can't it
> be caught at run time? My best get at explaining this is by what I wrote
> above, i.e. it's a consequence of Typed Racket's more refined static
> notion of type. In certain other systems this wouldn't be possible,
> because the static checker would only catch (albeit earlier) errors that
> would be caught at run time, so there would be no basis to call the
> result "wrong".

No. A buggy program can produce a value in Erasure where in Natural or
Transient it produces a run time type error. Greenman et al. [10] prove
formally this proposition but it is easy to construct an example that
demonstrates the issue. For instance if the client in Figure 1 just
formats the result of 'crypto-pack' to a string and returns it, then the
client will terminate without an error but its string result will be
wrong; it will be a string representation of a hash rather than a list.
The reason is that the `format` primitive of Racket does not care what its
argument is.

> line 424: "fully typed correct programs" -- presumably your method could
> also work with not-yet-fully-typed correct programs, just not ranging
> over the entire lattice in those cases. I was wondering whether that
> might give different/interesting results.

Yes. In fact, we already do this for programs that depend on library code.

> line 517: the difference between #1 and #2 here again relies on the
> surprising property (to the unfamiliar) that Typed Racket has a notion
> of "type-level mistake" that doesn't surface under erasure (as an
> exception or whatever) but also is not a false positive (i.e. you're not
> talking about conservativeness of static checking). Line 538's
> "unavoidable" claim is probably also true only in such a context.

Yes please see our answers to your questions about Erasure, line 34 and
line 313.

> line 756: "checked the value's type.." -- and the check passed!?

Indeed, Transient checks only the top-level type constructor of a
value. We will add a reminder of section 3's discussion of Transient
checks to the prose here.

> In the title, "evaluating blame" reads oddly. It is ambiguous, because
> "evaluate" sometimes means to compute a result. A fuller phrase, like
> "evaluating the usefulness of blame tracking...", would probably be
> worth the words.

> The abstract doesn't say much about the work. It would be better written
> for experts to quickly gather an overview what the paper contributes.

> Conversely, the main body of the paper skimps a bit on background. It
> never explicitly covers what "blame" means and how it works in practice.
> Similarly, it repeatedly talks about "impedance mismatches" without
> defining them. Perhaps this phrase is now standard in the gradual typing
> literature, which I haven't kept up with (hence my Z expertise). But I
> did read the Wadler/Findler paper carefully at the time. It did not talk
> about impedance mismatches. In any case, it is a fuzzy metaphor...
> please say exactly what it means here.

> line 119: "homo economicus" needs glossing or a reference

> line 94: "forego" => "forgo"

> line 125: was wondering whether "impedance mismatch" just means
> "feasible run-time type error".

> line 148: last sentence appears to contradict the preceding paragraph,
> and is left hanging oddly. Instead, make it the start of a new
> paragraph.

> line 160: "disparity" is a strange word here. "diversity"?

> line 170: "Otherwise, results from..." -- the point was already clear

> line 174: ... instead of just saying what you built on, first say what
> you did! It's really not clear at this point.

> line 199: "blame set as another form of a stack trace" -- yes. I was hoping to
> see a clearer example of debugging with and without blame, i.e. something
> making explicit the similarities and differences between having blame info and
> having only a stack trace at the error site.

> line 211: "languages exceptions" typo

> line 268: this explanation of the workings of proxies seemed overwrought

> line 283: "rewrites typed modules to inline checks" -- so a module gets
> turned into a check? Clearly not, but that's how it reads

> line 299: "crosses" => "crossings" (probably)

> lien 352: "three interpretations" -- what are they? I don't see them in
> the figure.

> line 380: from this I inferred that "migratory typing" means "gradual
> typing applied at modulewise granularity". Assuming that's correct, it's
> worth saying directly.

> line 421: "type mistake" -- does this mean "feasible run-time type
> error"?

> line 433: "without loss of diversity" -- this is glib. Clearly diversity
> is lost; just claim that what remains is still diverse enough.

> line 497: "truthiness" needs explaining

> line 519: "at least three" -- better to claim this is only two here,
> then explain later that the driver doesn't count. Mentioning "three"
> up-front just raises an unnecessary question in the reader's mind.

> line 634: earlier I had inferred that the distinction between
> "migratory" and "gradual" was more than just preferring one word; this
> seems to contradict that

> line 640: what makes it "concise"?

> line 673: "programmers runs" typo

> line 727: "dubbed" -- reads oddly. Maybe italicise the "location", but
> it seems overkill. The phrase is pretty self-suggesting as it is.

> line 752: "added to the blame set first" -- should it be a blame list,
> then?

> line 936: no need for hyphen after the adverb

> line 966: "more useful C" -- missing "than"

> line 982: don't think these percentages deserve 3 significant figures

> line 1304: "Problem is" -- missing "The". Also, no need to italicise the
> next sentence... it is really not that deep or surprising.

Thanks for all these detailed feedback points. We will use them to improve
the prose.