### 2021-02-23 writing/mutate-benchmarks-figure.tex
factor out figures

\begin{figure}
\begin{tabular}{p{1.75cm} | p{6cm} }
  {\bf  name} & {\bf description (author)}  \\

\hline
  \texttt{acquire} & Object-oriented board game implementation (M. Felleisen)  \\%[1em]

\hline
  \texttt{gregor} & Utilities for calendar dates (J. Zeppieri) \\%[1em]

\hline
  \texttt{kcfa} & Functional implementation of 2CFA for the lambda calculus (M. Might) \\%[1em]

\hline
  \texttt{quadT} & Converter from S-expression source code to PDF format (M. Butterick)\\%[1em]

\hline
  \texttt{quadU} & Converter from S-expression source code to PDF format  (B. Greenman) \\%[1em]

\hline
  \texttt{snake} & Functional implementation of the  Snake video game (D. Van Horn) \\%[1em]

\hline
  \texttt{suffixtree} & Algorithm for common longest subsequences between strings. (D. Yoo) \\%[1em]

\hline
  \texttt{synth} & Converter of notes and drum beats to WAV (V. St-Amour \& N. Toronto) \\%[1em]

\hline
  \texttt{take5} & Mixin-based card game simulator (M.Felleisen)  \\%[1em]

\hline
  \texttt{tetris} & Functional implementation of Tetris (D. Van Horn) \\%[1em]

\end{tabular}
  \caption{Benchmarks summary.}
  \label{table:benchmark-descriptions}
\end{figure}