### 2021-02-23 writing/mutate.tex

%% -----------------------------------------------------------------------------

As section~\ref{sec:challenges} explains, the first challenge is to
generate a representative mistakes in representative software. This
section explains the specific choices made for the first application
of the method: the source and nature of the chosen benchmarks
(sec.~\ref{sub:mutate-benchmarks}), the newly developed mutators
(sec.~\ref{sub:mutate-mutators}), and why they are suitable for this
experiment (sec.~\ref{sub:mutate-interesting}).

%% -----------------------------------------------------------------------------
\def\sub#1#2{\subsection{#2} \label{sub:mutate-#1} \input{mutate-#1.tex}}

\input{mutate-benchmarks-figure.tex}
\sub{benchmarks} {Which Software to Use As Benchmarks}
\sub{mutators}   {How to Mutate the Benchmarks} 
\input{mutate-mutators-figure.tex}
\sub{interesting}{Are These Mutators Interesting}