### 2021-03-01 writing/mutate.tex
re-organize and re-write secs. 1 thru 6

%% -----------------------------------------------------------------------------

Putting a rational programmer to work means generating many mutants and turning
those into debugging scenarios. Since existing mutators do not generate useful
mutants, the first step is to develop new mutators (see sec.~\ref{sub:mutate-mutators})
and to validate their suitability for this application (see sec.~\ref{sub:mutate-interesting}).

% As section~\ref{sec:challenges} explains, the first challenge is to
% generate a representative mistakes in representative software. This
% section explains the specific choices made for the first application
% of the method: the source and nature of the chosen benchmarks
% (sec.~\ref{sub:mutate-benchmarks}), the newly developed mutators
% (sec.~\ref{sub:mutate-mutators}), and why they are suitable for this
% experiment (sec.~\ref{sub:mutate-interesting}).

%% -----------------------------------------------------------------------------
\def\sub#1#2{\subsection{#2} \label{sub:mutate-#1} \input{mutate-#1.tex}}

\input{mutate-mutators-figure.tex}
\sub{mutators}   {How to Mutate Software} 
\sub{interesting}{Are These Mutators Interesting}