### 2021-03-01 writing/mutate-mutators.tex
adjust prose

% -----------------------------------------------------------------------------

A {\it mutator\/} performs a localized syntactic change to a code base. The
result is a {\em mutant\/}.

For the evaluation of a blame strategy, mutators must produce type-level
mistakes that the run-time checks of gradual typing systems or the safety
checks of the underlying language can detect. Once detected, the rational
programmer should be able to locate the mistake by gradually adding types to
blamed modules.

Figure~\ref{table:mutation-ops} describes 16 mutators that satisfy this
requirement.  For each mutator, the figure provides a short description and an
example. As the explanation indicates, some specialize or
generalize~\citet{lksfd-popl-2020}'s mutators, which in turn are borrowed from
the vast literature on mutation testing~\cite{jia2011analysis}.\footnote{Many
of~\citet{lksfd-popl-2020} mutators do not satisfy the constraints. For
example, the relational mutator does not lead to type-level errors that Typed
Racket's type system can detect, so it has been eliminated. Similarly, the
original arithmetic mutator does not reliably lead to type errors, so it has
been adapted to only make operator swaps that can affect the type of the
result. Changing a `/` to `*` does not affect the type of an expression, but
changing `*` to `/` may, because division by zero is detectable by Typed
Racket's type system.}  Only two are directly inherited; many mutators are
brand new. For the latter the authors relied on their decades-long experience
of making type-level mistakes in Typed Racket, some of which take a non-trivial
effort to debug.

Most of the mutators are self-explanatory.  The first six apply to all
gradually typed languages; the next six to those that include classes and
objects. The last four target distinguishing features of Typed Racket,
specifically is sophisticated type system. In particular, {\tt arithmetic}
replaces a {\tt +} with a {\tt -} in an attempt to change the type of the
arithmetic expression; {\tt +}'s result is a {\tt Positive-Integer} when all
arguments are positive integers, while {\tt -} yields {\tt
Integer}~\cite{stathff-padl-12}. Similarly, the other three also aim to usurp
Typed Racket's occurrence type system.