### 2021-03-01 writing/rational-stuff.tex
stuff

%% -----------------------------------------------------------------------------

\citet{tfgnvf-popl-2016} describe the set of all possible type migrations with a
lattice. Formally, a program $\system$ is a set of modules $\set{\component}$.
Let a configuration $\conf$ of $\system$ be the subset of $\set{\component}$ that
comes with (possibly type-incorrect) annotations. These configurations of
$\system$ are ordered by the subset relation and form a lattice
$\lattice{\system}$ with $2^{\size{\system}}$ elements. The bottom scenario of
$\lattice{\system}$ is $\emptyset$, the top one $\system$ itself. In between are
the mixed-typed scenarios.

Applying a mutator to module $\component^*$ of $\system$ acts like a homomorphism
on the generated lattice. The two lattices differ only in the mutant of
$\component^*$. Given this formulation, debugging scenarios are those points in
this new lattice that do not contain the mutant; at all other points, the type
checker points out the type-level mistake in the mutant of $\component^*$.

The actions of the rational programmer create an ascending chain---dubbed a
\emph{trail}---in $\lattice{\system}$.  The starting scenario $\conf_0$ of a
trail---also referred to as the \emph{root} (of the trail)---is the initial
interesting debugging scenario.  Starting from such a root, the rational
programmer steps along a trail. If the scenario type-checks, the programmers runs
the program until it raises a run time error.  The rational programmer uses the
information in the error to decide which component(s) to equip with types. This
choice constructs scenario $\conf_{i+1}$ from $\conf_{i}$ and thus creates a
trail. The trail ends when programmer reaches a scenarios that contain the mutant
module, because the type checker rejects it outright. At this point, the
programmer has identified the source of the impedance mismatch.