### 2021-03-01 writing/results-benchmarks-figure.tex
name files properly

\begin{figure}
\begin{tabular}{p{1.75cm} | p{4.2cm} | p{2.2cm} | p{2.2cm}}
  {\bf  name} & {\bf description (author)} & {\bf lines of code} & {\bf
  components} \\

\hline
  \texttt{acquire} & Object-oriented board game implementation (M.
  Felleisen) & 1941 & 9 \\%[1em]

\hline
  \texttt{gregor} & Utilities for calendar dates (J. Zeppieri) & 2336 & 13\\%[1em]

\hline
  \texttt{kcfa} & Functional implementation of 2CFA for the lambda
  calculus (M. Might) & 328 & 7\\%[1em]

\hline
  \texttt{quadT} & Converter from S-expression source code to PDF format
  (M. Butterick) & 7396 & 14\\%[1em]

\hline
  \texttt{quadU} & Converter from S-expression source code to PDF format
  (B. Greenman) & 7282 & 14 \\%[1em]

\hline
  \texttt{snake} & Functional implementation of the  Snake video game (D.
  Van Horn) & 182 & 8 \\%[1em]

\hline
  \texttt{suffixtree} & Algorithm for common longest subsequences between
  strings. (D. Yoo) & 1500 & 6 \\%[1em]

\hline
  \texttt{synth} & Converter of notes and drum beats to WAV (V. St-Amour
  \& N. Toronto) & 871 & 10 \\%[1em]

\hline
  \texttt{take5} & Mixin-based card game simulator (M.Felleisen)  & 465 &
  8\\%[1em]

\hline
  \texttt{tetris} & Functional implementation of Tetris (D. Van Horn) &
  280 & 9 \\%[1em]

\end{tabular}
  \caption{Benchmarks summary.}
  \label{table:benchmark-descriptions}
\end{figure}