### 2021-03-01 writing/results.tex
cleaning up prose

%% -----------------------------------------------------------------------------

At this point all ingredients are in place to describe the first application of
the evaluation method. Two things are needed: (1) the description of the
experimental setup, including the experimental questions; and (2) a set of
representative benchmark programs. The first two subsections address these
issues in order. The last one presents the observed results.

\def\rsub#1#2{\subsection{#2} \label{sub:#1} \input{results-#1.tex}}

\rsub{experiment}{The Experimental Process} 
\input{results-benchmarks-figure.tex}
\rsub{benchmarks}{The Experimental Benchmarks}
\rsub{observed}  {The Observed Results}