### 2021-03-01 writing/sample.tex

%% -----------------------------------------------------------------------------

%  from section~\ref{sec:mutate} (96,515,744)

As is, the chosen mutators generate a huge number of debugging scenarios,
approximately 1,000,000 for a suitable benchmark suite such as the one used for
the first experiment (see~sec.~\ref{sub:results-benchmarks}). This number of
scenarios is far too large to even identify the interesting ones among them.
Furthermore this population is heterogeneous; different scenarios come from
different mutants and different mutants are the result of a different mutator
and benchmark pair.

To make the experiment computationally feasible and statistically precise,
stratified uniform random sampling with three levels of stratification becomes
necessary.  The first two levels group mutants first by benchmark and then by
mutator.  The set of randomly select from the mutants of each benchmark is
checked to make sure that, per benchmark, the sample reflects the diversity of
mutants with respect to the mutators that generated them.  Specifically, the
process samples 80 mutants with interesting scenarios per benchmark,
evenly-distributed across all of the mutators that contribute mutants for the
benchmark.  At the third level, the sampling process draws 96 random interesting
scenarios from each configuration lattice with replacement.

%% This looks like a repeat of the above, made concrete for the benchmark set

% In summary, the sampling process takes advantage of ancillary information to
% deliver a sufficiently large representative set of scenarios.  Concretely, we
% test each mode of the rational programmer on 72,192 randomly selected
% interesting debugging scenarios spread across both benchmarks and mutators.
% Furthermore, stratified random sampling allows us to derive precise conclusions
% about the proportion of the whole population that satisfies a given trait from
% the sample despite the population's heterogeneity.  That property is exactly
% what we need to project comparisons about the usefulness of various modes of the
% rational programmer from the sample to the set of all interesting debugging
% scenarios.

%% we use the following algorithm to select the 80 mutants evenly distributed across all of the mutators that do produce mutants of a benchmark:
%% 1. Treat each mutator as a bucket containing the mutants produced by that mutator
%% 2. Let the target-sample-size be 80
%% 3. Let the bucket-sample-size be target-sample-size/number-of-buckets
%% 4. For each bucket, do
%%    1. Randomly remove min(bucket-sample-size, mutants-in(bucket)) mutants from the bucket
%%    2. Set target-sample-size = target-sample-size - (the number of mutants removed)
%% 5. If target-sample-size is empty, done; else, go back to 3.