### 2021-03-02 writing/mutate-benchmarks-figure.tex
editiroal voice thru 6

\begin{figure*}
\begin{tabular}{p{1.4cm} | p{6.9cm} | p{2.0cm} | r | r}
             {\bf  name} & {\bf description} & {\bf author} & {\bf loc} & {\bf mod.} \\ \hline
%% ---------------------------------------------------------------------------------------------------------------

\texttt{acquire} & object-oriented board game implementation      & M. Felleisen & 1941 & 9 \\ \hline

\texttt{gregor}  & utilities for calendar dates                   & J. Zeppieri  & 2336 & 13\\ \hline

\texttt{kcfa}    & functional implementation of 2CFA for LC       & M. Might & 328 & 7\\ \hline

\texttt{quadT}   & converter from S-expression source code to PDF & M. Butterick & 7396 & 14\\ \hline

\texttt{quadU}   & converter from S-expression source code to PDF & B. Greenman  & 7282 & 14 \\ \hline

\texttt{snake}   & functional implementation of the Snake game    & D. Van Horn & 182 & 8 \\ \hline

\texttt{synth}  & converter of notes and drum beats to WAV & V. St-Amour & 871 & 10 \\ \hline

\texttt{take5} & Mixin-based card game simulator & M.Felleisen & 465 & 8\\ \hline

\texttt{tetris} & Functional implementation of Tetris & D. Van Horn &   280 & 9 \\ \hline

\texttt{suffix\-tree} & algorithm for common longest subsequences between strings. & D. Yoo & 1500 & 6 \\

\end{tabular}

\caption{Benchmarks summary.} \label{table:benchmark-descriptions}

\end{figure*}