### 2021-03-02 writing/why-rational-methodology.tex
methodlogy

%% -----------------------------------------------------------------------------

Every time the rational programmer debugs a program, it is validation for
programming language researchers. It shows how their slogans and theorems help
programmers connect the dots. The example of blame assignment mechanisms makes
this point clearly. A blame-assignment mechanism provides information that,
according to programming language research, points toward the source of the
problem.

By contrast, when the rational programmer fails, it questions programming
language research. Specifically, it is indicative of the limited predictive
power of programming language theory with respect to the use of languages in
practice. Put differently, it suggests flaws in language design due to
misleading predictions.

In this way, the idea of a rational programmer implies an entire methodology for
evaluating the design of programming languages. At this point, this study of
methods supplies many questions whose answers might point to suitable evaluation
methods:

\begin{enumerate}

\item Does a language design provide information that guides the actions of a
 rational programmer?

\item Does the underlying theory suggest actions to the rational programmer?

\item Can this guidance be formulated as an algorithm on the program and the extra
information?

\item Does the underlying theory lead to a hypothesis about the effects of
 these actions?

\item Can this hypothesis be tested with a large-scale automated experiment?

\end{enumerate}

Here is how the answers to these methodological questions imagines an evaluation
method for blame-assignment mechanisms:

\begin{enumerate}

\item The design of blame-assignment mechanisms explicitly advertises the blame
information as helpful for debugging impedance mismatches.

\item The error messages of blame-assignment mechanisms include suspect
violations at the boundary of typed and untyped code fragments. The
Wadler-Findler slogan suggests the problem is a lack of types, meaning the
addition of types to the untyped fragment is the most natural action.

\item The step-by-step construction of paths based on error messages from
gradual-typing checks is clearly amenable to implementation, modulo the
ascription of types to modules.

\item The theory theories implies that blame assignment constrains the search
space that a developer must inspect to find the problem.

\end{enumerate}
Sadly, seeing an algorithm is different from using the algorithm to conduct a
data-gathering experiments. The specific challenges are spelled out in the next
section, and the following three sections explains ways of overcoming them.