### 2021-03-03 writing/mutate.tex

%% -----------------------------------------------------------------------------

Putting the rational programmer to work means generating many mutants and turning
those into debugging scenarios. The process must start with a suitable
collection of representative programs (see sec.~\ref{sub:mutate-benchmarks}).
Since existing mutators do not generate useful mutants, the first step is to
develop new mutators (see sec.~\ref{sub:mutate-mutators}) and to validate their
suitability for this application (see sec.~\ref{sub:mutate-interesting}).

% As section~\ref{sec:challenges} explains, the first challenge is to
% generate a representative mistakes in representative software. This
% section explains the specific choices made for the first application
% of the method: the source and nature of the chosen benchmarks
% (sec.~\ref{sub:mutate-benchmarks}), the newly developed mutators
% (sec.~\ref{sub:mutate-mutators}), and why they are suitable for this
% experiment (sec.~\ref{sub:mutate-interesting}).

%% -----------------------------------------------------------------------------
\def\sub#1#2{\subsection{#2} \label{sub:mutate-#1} \input{mutate-#1.tex}}

\input{mutate-benchmarks-figure.tex}
\sub{benchmarks}{The Experimental Benchmarks}
\input{mutate-mutators-figure.tex}
\sub{mutators}   {How to Mutate Software} 
\sub{interesting}{Are These Mutators Interesting}