### 2021-03-03 writing/rational-stuff.tex

%% -----------------------------------------------------------------------------

\citet{tfgnvf-popl-2016} describe the set of all possible type migrations with a
lattice. Formally, a program $\system$ is a set of modules $\set{\component}$.
Let a configuration $\conf$ of $\system$ be the subset of $\set{\component}$
that comes with type annotations. These configurations of $\system$ are ordered
by the subset relation and form a lattice $\lattice{\system}$ with
$2^{\size{\system}}$ elements. The bottom scenario of $\lattice{\system}$ is
$\emptyset$, the top one $\system$ itself; in between are the mixed-typed
scenarios.

Applying a mutator to module $\component^*$ of $\system$ acts like a homomorphism
on the generated lattice. The two lattices differ only in the mutated
$\component^*$. Given this formulation, debugging scenarios are those points in
this new lattice that do not contain the mutated module; at all other points, the type
checker points out the type-level mistake in the mutated $\component^*$.

The actions of the rational programmer create an ascending chain---dubbed a
\emph{trail}---in $\lattice{\system}$.  The starting scenario $\conf_0$ of a
trail---also referred to as the \emph{root} (of the trail)---is the initial
debugging scenario.  Starting from such a root, the rational
programmer steps along a trail. If the (program for this) scenario type-checks,
the rational programmer runs the program until it raises a run-time error.  The rational programmer
uses the information in the error to decide which module to equip with
types. This choice constructs scenario $\conf_{i+1}$ from $\conf_{i}$ and thus
creates a trail. The trail's construction ends when it reaches a scenario that
contains the mutated module because the type checker rejects its typed version
outright.  At this point, the source of the impedance mismatch is identified.