### 2021-03-04 writing/sample.tex

%% -----------------------------------------------------------------------------

%  from section~\ref{sec:mutate} (96,515,744)

As is, the chosen mutators generate approximately one million debugging
scenarios for the chosen benchmarks. This number of scenarios is far too large
to even identify the interesting ones among them.  Furthermore this population
is heterogeneous; scenarios come from different mutants, and the mutants are the
result of different mutators applied to each benchmark.

To become computationally feasible and statistically sound, the experiment employs
stratified uniform random sampling with three levels of stratification.
The first two levels group mutants first by benchmark and then by
mutator, making sure that the sample within each benchmark reflects the diversity of
mutants with respect to the mutators that generated them.  Specifically, the
experiment samples 80 interesting mutants per benchmark,
evenly-distributed across all of the mutators that contribute mutants for the
benchmark. Some benchmarks have less than  80 mutants with interesting scenarios,
in which case the only choice is to include them all.
The result is a total of 756 interesting mutants across all benchmarks.
Finally, the third level of sampling draws 96 random interesting debugging
scenarios from each configuration lattice with replacement. The
final sample thus consists of 72,192 interesting scenarios.

%% MF: I don't get this number ^^^^
% > (* 756 96)
% 72576