### 2021-03-05 writing/mutate-benchmarks.tex

%% -----------------------------------------------------------------------------

The benchmark programs for a rational-programmer experiment must 
\begin{enumerate}

\item vary in size, complexity and purpose;

\item have a decent number of modules and a variety of module dependency graphs
 (because mixing of typed and untyped code in Typed Racket takes place at the
 module level);

\item take advantage of the variety of typing features of a gradually typed
 language; but

\item fail and contain type-level mistakes. 
\end{enumerate}
No such benchmark suite exists.  Worse, no benchmark suite of
gradually typed programs with type-level mistakes exists at
all.\footnote{\citet{cc-oopsla-20} created a collection of Reticulated
Python programs to evaluate their technique of fixing mistakes in type
annotations. Their collection does not come with type-level mistakes
in the code itself.}

What exists is \citet{gtnffvf-jfp-2019}'s collection of Typed Racket programs
for systematically measuring the implementation's performance. The benchmark
suite consists of fully typed, correct programs, written by a number of
different authors, who have maintained and evolved these programs over several
years. The programs range widely in size, complexity, purpose, origin and in
programming style. The latter is mostly a reference to their reliance on many
Typed Racket features: occurrence typing~\citep{tf-icfp-2010}, types for mutable
and immutable data structures~\citep{hpst-sfp-2010}, types for first-class
classes and objects~\citep{tsdtf-oopsla-2012}, and types for Racket's numeric
tower~\citep{stathff-padl-12}.

Hence, even if they do not come with type-level errors, they offer a diverse
starter suite, suitable for the careful introduction of type-level errors via
mutations. Further filtering the benchmark suite to keep the benchmarks with the
most sizeable and complex dependency graphs yields ten suitable programs while
preserving the suite's linguistic diversity. Figure~\ref{table:benchmark-descriptions} displays the selected
benchmark programs and describes their basic characteristics.