### 2021-04-22 writing/reviews/icfp21-rebuttal.md
Minor revisions of point-by-point replies

The first part of our response addresses big themes in the reviews;
the second part addresses all remaining concerns individually.

What do the results say about blame for gradual typing in practice?
================================================================================

The introduction says that the experimental results lead to the
following conclusions about blame _in practice_:

1. Blame helps locate impedance mismatches, though not immensely more than simple stacktraces, and
2. Natural blame helps more than Transient, and Transient more than Erasure's stacktraces, *but*
3. Natural blame is only marginally better than Transient in practice, and both not immensely better than Erasure's stacktraces.

The phrase "in practice" refers to the way that blame behaves in real
programs with an impedance mismatch, using real implementations of
each semantics.

In light of the data in figures 8, 9, and 10, the benefits of blame in
practice do not measure up to the expectations established by theory.
Concretely, blame does provide useful debugging information, but the
results suggest that in practice blame is only necessary to debug (as
compared with simple exceptions) in a small fraction of scenarios.

- The theory suggests that Natural's blame should be more precise than
  Transient because Natural precisely tracks all possible interactions
  while Transient approximates (see [11]: Greenman et al. OOPSLA'19).

- In practice, Natural's blame has utility comparable to Transient's.

The experiment in this paper reveals that mismatch, but it necessarily
considers only a limited view of the space of possible impedance
mismatch scenarios that might arise in reality.

Section 11 points to another space of scenarios that deserves
analysis: scenarios where mismatches arise due to mistakes in type
annotations rather than (or as well as) code.  Indeed, review B
astutely points out two more spaces worth exploring:

- programs where some components _cannot be typed_
- non-deterministic programs, which inject ephemeral bugs.

<!-- As another example, the paper only considers scenarios known to
be detectable by all three systems, but theory makes it clear that
Natural and Erasure can detect some scenarios that Erasure cannot. -->

<!-- How common are those scenarios for impedance mismatch bugs? -->
<!-- Does blame provide more benefit than simple exceptions in those scenarios? -->

Additional experimentation is necessary to develop an understanding of these different spaces.

The Rational Programmer vs Lazarek et al.
================================================================================

Lazarek et al. (POPL'20) provide a kernel insight, which this paper
develops into the novel idea of the Rational Programmer. It is the
first paper to explain the idea properly and use the phrase. We
therefore consider it its first contribution.

The key distinction is the idea of the rational programmer as an
entity to study and how it can be parameterized by modes to enable
cross-system comparison.

Replies to individual points
==================================================================

Review #45A
===========================================================================

> p5. one thing I wonder about is how this all hinges on the rational programmer not only taking the right action, but also taking it right? Is that exactly what the rational programmer idea presupposes? (I guess this reflects on your remark in Sec. 11 that one problem is with errors in the ascribed types themselves.).

Yes, you hit the nail on the head. 
One of the central parts of the rational programmer is that it fixes the decision of how to respond to a given scenario; 
so evaluating different ways of taking action boils down to defining and testing the corresponding rational programmers.

> When the rational programmer "fixes" a module and reruns, the implicit assumption seems to be that he runs the program in the exact same way. That would make the most sense to me, but you do not seem to come out and say so.

Yes you're right. Thanks, we will clarify this in the prose.

> It would help if you could explain how a mutator like negate-cond actually leads to a program
> that has more type-level mistakes than before. It does not seem to me this part of the code
> itself will now fail, but that changing the conditions makes certain previously excluded paths in the code feasible.

Thanks for pointing this out, we will expand the explanation and examples of mutators in the final paper.
In the specific cases of `negate-cond` and `force-cond`, these mutators lead to type-level mistakes when the program uses occurrence typing. 
For example, they could both lead to a mistake in a program like this:
```
(: use-int  (-> Integer Void))
(: use-bool (-> Boolean Void))
(: x        (U Integer Boolean))

(if (integer? x) (use-int x) (use-bool x))
```

> How do you arrive at the number 72192 in Section 7? It is not equal to 756 times 96.

Indeed, 756 is a typo: it should be 752.

Review #45B
===========================================================================

>  - Which blame assignment strategy is "good" and which is not "good"?

TODO - Christos will do it.

>  - At first glance, the conclusion in lines 55-56 (starting with "Second, ...")
>    seems to contradict that in line 99-100 (starting with "neither is ...").  It would
>    be nicer to address them in a clearer manner.

Thanks for the feedback, we will rephrase the conclusion.

> - The paper often mixes gradual typing and migratory typing.  It is unclear how
>   it distinguishes them.

The paper uses the two terms interchangeably after the introduction, because the distinction between the two does not change the evaluation method.
Specifically, in the gradual typing setting the components are smaller than whole modules, and there are more possible intermediate types (incorporating Dyn) between untyped and fully-typed, both of which just end up making the configuration lattice larger.

> - In principle, Natural blame should always point out the faulty components
>   by the Wadler-Findler slogan.  However, the experiment shows it is not the
>   case.  Why?  Is it the same reason as in Lazarek et al.?  (Perhaps it is due
>   to a gap between theory and practice, but exposing a reason is crucial.)

(Combining this question with the related one from below next.)

> - Are failing Natural blame trails produced even for programs with impedance
>   mismatch?  (This question is related to the above issue with Natural blame.)

We agree that the paper needs a bit more explanation to clarify how Natural blame can have failing
scenarios, we will update it in the final paper.  In short, Natural blame
trails can by stymied by unhelpful run-time errors.

Because debugging scenarios are only partially typed, some scenarios
produce an error from the runtime (e.g. `+` receiving a non-number)
before any of Natural's run-time type checks can discover a problem.
Run-time errors do not carry blame, only stacktraces, so the only way
to proceed (regardless of mode) is by using the stack information.  If
the stack proves unhelpful, then the trail fails.

> - It is not fully explained how the given mutators change programs.  For
>   example, I cannot completely predict changes by the mutators deletion and
>   class:super.

Thanks for the comment, we will expand the explanation and examples of mutators in the final paper.
In the specific case of `deletion`, the most common place that this mutator applies
(beyond the literal `begin` expressions the example illustrates)
are so-called "implicit begins": places where
sequences are allowed implicitly by enclosing forms like `define`
or `cond`.
Here are two more examples of `deletion` mutations for these situations:
```
(define (do-something) (step-1!) (step-2!))
~>
(define (do-something) (step-1!))

(cond [(do-something?) (step-1!) (step-2!)]
      [else (void)])
~>
(cond [(do-something?) (step-1!)]
      [else (void)])
```

In the case of `class:super`, the mutator simply removes calls to `super-new`, causing them to no longer instantiate the superclass.

> - The paper says that the experiment uses 72,192 sampled scenarios (line 802).
>   However, the total number of mutants (i.e., configuration lattices) is 756
>   and, from each lattice, 96 scenarios are sampled.  Thus, I guess 756*96 =
>   72576 scenarios are used.  Why does this mismatch happen?

Indeed, 756 is a typo: it should be 752.

> - The experiment samples multiple scenarios from a configuration lattice (line
>   800).  Does this mean the debugging process may start with a mix-typed program
>   and may not with a fully untyped one?  If so, why does the paper take such an
>   approach?

The bottom of the configuration lattice is completely untyped, so that
is a possible starting scenario to be sampled.

> - The paper often says that a blame system is (un)sound, but it is difficult for
>   me to identify what it precisely means.

The term refers to the soundness defined by [11] (Greenman et
al. OOPSLA'19), but we agree that for clarity the paper should define
it in addition to the reference.  Intuitively, a blame system is sound
if it can only blame components that have had control of the value
witnessing a violation.  In other words, the system cannot blame
arbitrary components unrelated to the witness of a contract violation.

> - The paragraph starting at line 1122 is quite difficult to understand for me.
>   [...]
>   - What does the paper mean by "incomplete population of the blame map"?
>   - To understand the idea on the improvements of Transit, more explanations
>     on the usage of blame maps in Transient would be needed.

We will improve the prose in this section to make it clearer, thank you for pointing it out.
The paper is referring here to the incomplete population of the blame map described by Greenman et al. ([11]: OOPSLA'19) .
For more details, there is a parallel submission by the same set of authors describing the implementation of Transient in depth.

>   - What "sophisticated typing features" are considered (line 589)?  It would be
>     crucial to confirm whether the proposed mutators are enough.

Broadly speaking, this phrase refers to features beyond basic functions and
primitive data. In terms of this paper's evaluation, the benchmarks employ many
of Typed Racket's sophisticated features (with those listed on L641 being the
most notable).

We absolutely agree that it is crucial to analyze the mutators in light of these
features; perhaps adding a breakdown of features per benchmark to the analysis
of section 6.3 would help clarify this?

>   - The benchmarks are selected (line 647), but why?  The GPT benchmark suite of
>     Racket provides more examples.

The benchmarks in figure 5 are the ten with the largest dependency graphs of the GTP suite.
We will rephrase L646-648 to make this clearer.

> - It seems that the experiment implicitly supposes a few assumptions on
>   benchmarks.  First, the original benchmarks must be fully typed.  Thus more
>   experiments on mix-typed programs where some components cannot be typed may be
>   needed.  Second, the benchmarks are supposed to be deterministic, which I find
>   from the definitions of trails.  If it is the case, while I do not think these
>   restrictions have to be lifted in the paper, it would be nice to expose them.

You are absolutely right, we will clarify these assumptions about the
benchmarks in the paper. They are inherited from the several
publications on the gradual benchmark suite.

> - As an alternative of Transient first and last blame, perhaps determining
>   blamed modules by voting (i.e., blaming a module that has been most often
>   added to the blame map) might be promising (if the blame map has a large
>   population).  Have the authors considered it or another approach for Transient
>   blame?

We have since thought of a few alternative strategies, but not that
one: good idea!  Ultimately, Transient does not offer any
interpretation of its blame, so the paper chooses first and last as
two obvious reasonable choices.

> - Perhaps it is valuable for followers to share the experience on developing
>   mutators that are not interesting, .e.g., in the supplementary material.

We agree and would be happy to do so.

> - Why doesn't the paper adopt the author-year citation style?

This was a mistake typesetting the paper, we will fix it.

> - Page 7: How is the question (5) answered?

Thanks for pointing out that this is confusing.  As it is, the rest of
the paper is the implicit answer to (5), but we will clarify the
prose.

> - L356 "despite advertisements for the opposite":  I cannot find what this means.

This is meant to convey that although Lazarek et al. aim to evaluate
blame, they fail to adequately account for confounding factors.  We
will reword for clarity.

> - L370 "the latter must represent":  What does the "latter" specify?

It refers to impedance mismatches from the end of the preceding
sentence; we will clarify the wording here.

> - L424 "Let a configuration s of P":  Are configurations the same as scenarios?

Scenarios are a specific kind of configuration (that doesn't type the
buggy module).  We will adjust these two paragraphs to make this more
clear.

> - L431 "type-level mistake":  Is it the same as a impedance mismatch?

A type-level mistake is a mistake that causes an impedance mismatch at
a boundary between typed and untyped code.  We will clarify this point
in the prose.

> - L556: It would be helpful to describe how to extend trails and how to
>   determine if there is no scenario to be added.  (This comment is also related
>   to the issue with Natural blame).

Thanks, we will expand on this in the prose.

> - L557: "if it does" --> "if it does not"?

You're right, thanks, we will fix it.

> - L685 "e.g., changing '*' to '/'":  Is this an example of the mutator that
>   "does not reliably lead to type errors"?

This is a refinement of the arithmetic mutator that does lead to type
errors, because given two integers `*` always produces an integer, but
`/` may produce a rational, which have different types in Typed
Racket.  We will clarify this comment in the footnote.

> - L785 "the interesting standard guided countless iterations":  I fail to find
>   what this intends.

The measure of interesting mutants provided a standard by which we
could judge potential mutators and iteratively develop an effective
set of them.  We will make this comment clearer.

> - What is the total number of failing trails in Transient?

3594 out of 72192 trails fail for transient last blame, and 3591 for
first.

> - L1099 "as behavioral economics has shown more recently":  Is there a reference
>   to be cited?

The best reference we might use here is Kahnemann's popular book entitled "Thinking Fast and Slow", which explains the problem well.
Notably with respect to this work, the book criticizes classical (mathematical) economic theory for assuming that people act rationally all of the time, which demonstrates how our work is complementary to human studies; both are necessary.

> - L1104 "deviating is a mistake":  Why?

Our long experience of ignoring blame as opposed to heeding it
suggests that ignoring it is usually unwise.

> - L1154 "in the "fully typed" benchmarks":  Are these benchmarks on Python?

The slowdowns the paper quotes here are from Python benchmarks.

> - L1161 "the simplest benchmark":  Is this on Racket?  Which benchmark
>   does it intend?

This paragraph refers to a Python translation of the Racket `Sieve`
benchmark; it is confusingly worded, we will adjust the phrasing.

> - L118: "(at the mid-level on the left)[] imports" (the comma is removed)
> - L332 "soundness mechanisms":  What they mean is somewhat unclear.  Please
>   consider clarifying.
> - L347 "the rational programmer's effort": This was unclear at the first
>   reading.  It would be nice to briefly explain it here.
> - L451: "denotes the module [of P] that s blames"
> - L970: "the experiment provides [an] evidence that"
> - L1198 "But just because...":  It is difficult for me to find what this
>   sentence wants to say.  Please consider rephrasing.

Thanks for these comments, we will incorporate them to improve the prose.