### 2021-04-22 writing/reviews/icfp21-rebuttal.md
rebuttal revisions

The first part of our response addresses one selected theme from each review;
the second part addresses all remaining comments and questions individually.

What the experimental results say about the theory vs the practice of blame
================================================================================

Theoretical work that compares different checking and blame strategies for
gradual types (Greenman et al. ICFP 2018 and OOPSLA 2019) declares a clear
winner: Natural. Natural is the only semantics that is strongly type sound,
a complete monitor and, most importantly in the context of this paper, has
sound and complete blame assignment. The latter means that Natural blames
all and only those components that are relevant to a type impedance. In
contrast Transient may fail to blame some relevant components and may
blame irrelevant ones.

At first glance, the experiment of this paper seems to validate the
theoretical conclusions:

1. Natural's blame assignment is as good as the theory predicts; 
Natural blame helps locate impedance mismatches in almost all 
debugging scenarios. The only exceptions are (i) due to bugs in Typed
Racket and (ii) debugging scenarios that result in runtime errors with
unhelpful stacktraces from the underlying checks of Racket rather than 
Typed Racket's checks.

2. In general, Natural blame is better that Transient blame and both are
better than Erasure's stacktraces.

However, a closer inspection of the empirical results reveals that 
they paint a picture with gray zones that the black and white theoretical 
results are  not refined enough to capture:

1. There is a small but substantial number of debugging scenarios where 
Natural blame is not helpful but Transient blame is. These scenarios are
not all due to bugs in Typed Racket. Hence, the theory that deems natural
blame superior to that of Natural cannot explain them away.

2. While the theory claims that language designers should select Natural and
   its blames for maximum benefits, the results of the experiment indicate
   that Transient and Erasure are good enough for locating impedance
   mismatches in most cases. Hence, given the cost of Natural, the theory's answer
   fails to provide good enough guidance to language designers.

In sum the experiment in this paper reveals the mismatch between the theory and the practice
of blame. The latter refers to blame's actual utility in real
programs that run in a real implementation rather than artificial
examples in an idealized model.

As a result, there are two intertwined directions forward:

1. Revisiting the theory of blame to match the practice of blame.

2. Analyzing further the practice of blame beyond the scenarios this paper
   considers.

Section 11 points to another space of scenarios that deserves
analysis: scenarios where mismatches arise due to mistakes in type
annotations rather than (or as well as) code.  Indeed, review B
astutely points out two more spaces worth exploring:

- programs where some components _cannot be typed_
- non-deterministic programs, which inject ephemeral bugs.

What is the origin of the rational programmer
================================================================================

Lazarek et al. (POPL'20) provide the kernel insight that one can simulate
an idealized programmer on a large corpus of programs to test the behavior
of a language feature. This paper hones this kernel into an experimental
framework that allows the isolation of confounding factors and,
furthermore, the comparison between different semantics for a feature in
the form of different but comparable experimental modes. The paper
christens the experimental framework the Rational Programmer.

Replies to individual points
==================================================================

Review #45A
===========================================================================

> p5. one thing I wonder about is how this all hinges on the rational programmer not only taking the right action, but also taking it right? Is that exactly what the rational programmer idea presupposes? (I guess this reflects on your remark in Sec. 11 that one problem is with errors in the ascribed types themselves.).
Yes, you hit the nail on the head. 
One of the central parts of the rational programmer is that it fixes decision of how to respond to a given scenario; 
so evaluating different ways of taking action boils down to defining and testing the corresponding rational programmers.

> When the rational programmer "fixes" a module and reruns, the implicit assumption seems to be that he runs the program in the exact same way. That would make the most sense to me, but you do not seem to come out and say so.
Yes you're right. Thanks, we will clarify this in the prose.

> It would help if you could explain how a mutator like negate-cond actually leads to a program
> that has more type-level mistakes than before. It does not seem to me this part of the code
> itself will now fail, but that changing the conditions makes certain previously excluded paths in the code feasible.
In the specific cases of `negate-cond` and `force-cond`, these mutators lead to type-level mistakes when the program uses occurrence typing. 
For example, they could both lead to a mistake in a program like this:
```
(: use-int  (-> Integer Void))
(: use-bool (-> Boolean Void))
(: x        (U Integer Boolean))

(if (integer? x) (use-int x) (use-bool x))
```

> How do you arrive at the number 72192 in Section 7? It is not equal to 756 times 96.
Indeed, 756 is a typo: it should be 752.

Review #45B
===========================================================================

>  - Which blame assignment strategy is "good" and which is not "good"?
TODO

>  - At first glance, the conclusion in lines 55-56 (starting with "Second, ...")
>    seems to contradict that in line 99-100 (starting with "neither is ...").  It would
>    be nicer to address them in a clearer manner.
Thanks for the feedback, we will rephrase the conclusion.

> - The paper often mixes gradual typing and migratory typing.  It is unclear how
>   it distinguishes them.
TODO

> - It is not fully explained how the given mutators change programs.  For
>   example, I cannot completely predict changes by the mutators deletion and
>   class:super.
In the specific case of `deletion`, the other most common place that
>   this mutator applies are so-called "implicit begins": places where
>   sequences are allowed implicitly by enclosing forms like `define`
>   or `cond`. 
So here are two more examples of `deletion` mutations:
```
(define (do-something) (step-1!) (step-2!))
~>
(define (do-something) (step-1!))

(cond [(do-something?) (step-1!) (step-2!)]
      [else (void)])
~>
(cond [(do-something?) (step-1!)]
      [else (void)])
```

In the case of `class:super`, the mutator simply replaces any occurrence of `super-new` in the body of a class definition with `void`.

> - The paper says that the experiment uses 72,192 sampled scenarios (line 802).
>   However, the total number of mutants (i.e., configuration lattices) is 756
>   and, from each lattice, 96 scenarios are sampled.  Thus, I guess 756*96 =
>   72576 scenarios are used.  Why does this mismatch happen?

Indeed, 756 is a typo: it should be 752.

> - The experiment samples multiple scenarios from a configuration lattice (line
>   800).  Does this mean the debugging process may start with a mix-typed program
>   and may not with a fully untyped one?  If so, why does the paper take such an
>   approach?

The bottom of the configuration lattice is completely untyped, so that
is a possible starting scenario to be sampled.

> - The paper often says that a blame system is (un)sound, but it is difficult for
>   me to identify what it precisely means.

The term refers to the soundness defined by [11] (Greenman et
al. OOPSLA'19), but we agree that for clarity the paper should define
it in addition to the reference.  Intuitively, a blame system is sound
if it can only blame components that have had control of the value
witnessing a violation.  In other words, the system cannot blame
arbitrary components unrelated to the witness of a contract violation.

> - The paragraph starting at line 1122 is quite difficult to understand for me.
>   [...]
>   - What does the paper mean by "incomplete population of the blame map"?
>   - To understand the idea on the improvements of Transit, more explanations
>     on the usage of blame maps in Transient would be needed.
TODO

>   - What "sophisticated typing features" are considered (line 589)?  It would be
>     crucial to confirm whether the proposed mutators are enough.

The benchmarks employ many of Typed Racket's sophisticated features,
most notably (from L641) occurrence typing, types for mutable and
immutable data structures, types for first-class classes and objects,
and types for Racket's numeric tower.  We absolutely agree that it is
crucial to analyze the mutators in light of these features; perhaps
adding a breakdown of features by benchmark to section 6.3 would help
clarify this?

>   - The benchmarks are selected (line 647), but why?  The GPT benchmark suite of
>     Racket provides more examples.

The benchmarks in figure 5 are the ten largest dependency graphs of the GTP suite. 
We will rephrase L646-648 to make this clearer.

> - It seems that the experiment implicitly supposes a few assumptions on
>   benchmarks.  First, the original benchmarks must be fully typed.  Thus more
>   experiments on mix-typed programs where some components cannot be typed may be
>   needed.  Second, the benchmarks are supposed to be deterministic, which I find
>   from the definitions of trails.  If it is the case, while I do not think these
>   restrictions have to be lifted in the paper, it would be nice to expose them.

You are absolutely right, we will clarify these assumptions about the
benchmarks in the paper. They are inherited from the several
publications on the gradual benchmark suite.

> - As an alternative of Transient first and last blame, perhaps determining
>   blamed modules by voting (i.e., blaming a module that has been most often
>   added to the blame map) might be promising (if the blame map has a large
>   population).  Have the authors considered it or another approach for Transient
>   blame?

We have since thought of a few alternative strategies, but not that
one: good idea!  Ultimately, Transient does not offer any
interpretation of its blame, so the paper chooses first and last as
two obvious reasonable choices.

> - Perhaps it is valuable for followers to share the experience on developing
>   mutators that are not interesting, .e.g., in the supplementary material.

We agree and would be happy to do so.

> - Why doesn't the paper adopt the author-year citation style?

This was a mistake typesetting the paper, we will fix it.

> - Page 7: How is the question (5) answered?

Thanks for pointing out that this is confusing.  As it is, the rest of
the paper is the implicit answer to (5), but we will clarify the
prose.

> - L356 "despite advertisements for the opposite":  I cannot find what this means.

This is meant to convey that although Lazarek et al. aim to evaluate
blame, they fail to adequately account for confounding factors.  We
will reword for clarity.

> - L370 "the latter must represent":  What does the "latter" specify?

It refers to impedance mismatches from the end of the preceding
sentence; we will clarify the wording here.

> - L424 "Let a configuration s of P":  Are configurations the same as scenarios?

Scenarios are a specific kind of configuration (that doesn't type the
buggy module).  We will adjust these two paragraphs to make this more
clear.

> - L431 "type-level mistake":  Is it the same as a impedance mismatch?

A type-level mistake is a mistake that causes an impedance mismatch at
a boundary between typed and untyped code.  We will clarify this point
in the prose.

> - Are failing Natural blame trails produced even for programs with impedance
>   mismatch?  (This question is related to the above issue with Natural blame.)

Yes, please see the start of this response.

> - L556: It would be helpful to describe how to extend trails and how to
>   determine if there is no scenario to be added.  (This comment is also related
>   to the issue with Natural blame).

Thanks, we will expand on this in the prose.

> - L557: "if it does" --> "if it does not"?

You're right, thanks, we will fix it.

> - L685 "e.g., changing '*' to '/'":  Is this an example of the mutator that
>   "does not reliably lead to type errors"?

This is a refinement of the arithmetic mutator that does lead to type
errors, because given two integers `*` always produces an integer, but
`/` may produce a rational, which have different types in Typed
Racket.  We will clarify this comment in the footnote.

> - L785 "the interesting standard guided countless iterations":  I fail to find
>   what this intends.

The measure of interesting mutants provided a standard by which we
could judge potential mutators and iteratively develop an effective
set of them.  We will make this comment clearer.

> - What is the total number of failing trails in Transient?

3594 out of 72192 trails fail for transient last blame, and 3591 for
first.

> - L1099 "as behavioral economics has shown more recently":  Is there a reference
>   to be cited?

TODO

> - L1104 "deviating is a mistake":  Why?

Our long experience of ignoring blame as opposed to heeding it
suggests that ignoring it is usually unwise.

> - L1154 "in the "fully typed" benchmarks":  Are these benchmarks on Python?

The slowdowns the paper quotes here are from Python benchmarks.

> - L1161 "the simplest benchmark":  Is this on Racket?  Which benchmark
>   does it intend?

This paragraph refers to a Python translation of the Racket `Sieve`
benchmark; it is confusingly worded, we will adjust the phrasing.

> - L118: "(at the mid-level on the left)[] imports" (the comma is removed)
> - L332 "soundness mechanisms":  What they mean is somewhat unclear.  Please
>   consider clarifying.
> - L347 "the rational programmer's effort": This was unclear at the first
>   reading.  It would be nice to briefly explain it here.
> - L451: "denotes the module [of P] that s blames"
> - L970: "the experiment provides [an] evidence that"
> - L1198 "But just because...":  It is difficult for me to find what this
>   sentence wants to say.  Please consider rephrasing.

Thanks for these comments, we will incorporate them to improve the prose.