### 2021-05-17 writing/related.tex
edits to related work to fix monotoinc references discussion

%% -----------------------------------------------------------------------------

\citet{lksfd-popl-2020}'s human-out-of-the-loop empirical analysis of blame for
higher-order contracts is the inspiration for this paper.  While they do not
spell out the notion of the rational programmer, they present the basic
ingredients of the technique.

At a technical level, this work and~\citet{lksfd-popl-2020}'s differ in
many ways. The experiment presented here concerns three different modes of
gradual typing, theirs a single notion of contract checking.  It also
contributes the idea of creating three comparable rational programmers
(with several modes). Thus the method of this paper answers whether blame
adds value to a gradual type system and what gradual type system it
benefits the most; while~\citet{lksfd-popl-2020} only establish a
systematic relation between blame for contracts and behavioral bugs. The
type-level mutators represent another technical contribution of the work
presented here; as explained in section~\ref{sec:mutate}, almost none
of~\citet{lksfd-popl-2020}'s mutators is useful at the level of type
mistakes.

The literature on gradual typing presents many semantics beyond the three used
here and three additional blame strategies.  Pyret (pyret.org) assigns fixed-size
data types the Natural semantics and functions a Transient semantics. The
Forgetful~\citep{cl-icfp-2017} and the Amnesic~\citep{gfd-oopsla-2019} semantics
are similar to Transient but use wrappers instead of in-lined checks.
Nom~\citep{mt-oopsla-2017} and other \emph{concrete\/}
semantics~\citep{wnlov-popl-2010, rsfbv-popl-2015, rzv-ecoop-2015,
rat-oopsla-2017} assume that every value comes with a type tag and use tag
checks to supervise the interactions between typed and untyped code.  The semantics~\citep{tlt-popl-2019,
etg-icfp-19, tt-scp-20, tgt-popl-18, tt-sas-17} derived with the Abstracting
Gradual Typing technique~\citep{gct-popl-2016} are variants of Natural.
 Monotonic References~\citep{svctg-esop-2015, rsfbv-popl-2015, sfrbcsb-popl-2014, kas-pldi-2019} 
differ from Natural in the treatment of mutable data by
associating heap locations with a type and permitting only updates that 
make that type more precise.  
Among these representative semantics, only Amnesic, Nom and Monotonic References present innovative blame
strategies.  The experiment excludes the first because it is
merely a theoretical construction, the second because it imposes severe restrictions on
programmers and the third because it would require a complete
re-engineering of the Racket runtime. That said, the paper offers a
detailed chart that the creators of these blame strategies can transplant to their
linguistic setting and determine the value of their blame strategies.