### 2021-05-18 writing/why-rational.tex

%% -----------------------------------------------------------------------------

The general challenge of evaluating blame is a methodological one. Unlike most
current research on programming languages, the question seems to call for
empirical studies similar to those of the human-computer interaction research
area. At the same time, a significant result demands a large amount of
data. As~\citet{lksfd-popl-2020} recently demonstrated, the way around this
dilemma is to simulate a programmer algorithmically on a large set
of programming scenarios.

\input{why-rational-case.tex}

This paper generalizes Lazarek et. al.'s idea to {\em the rational programmer\/}.
Like~\citet{mill1874essays}'s {\it homo economicus\/}, the rational programmer approximates the
behavior of a software developer who reduces time spent on a task by exploiting the
available information. In the context of gradual typing, the rational programmer
has two pieces of information when an impedance mismatch signals exceptional
behavior: the error message and the state of the program. Hence, the most
rational procedure is to use the former to improve the latter. Specifically,
the rational programmer translates the Wadler-Findler slogan into a debugging
method, searching for the source of the impedance mismatch by adding type
annotations to some of the untyped parts of the program identified in the error.
If the type checker rejects an annotation derived from the context, the
rational programmer has found the source of the problem.
Otherwise, the rational programmer concludes that the just-annotated parts are not the problem
and re-runs the program;
the Wadler-Findler slogan implies the new version of the program must now blame
a different location for the problem, which the rational programmer can check in
the same fashion. 
Measuring this simulated
behavior on a large number of debugging scenarios yields data that
is similar to data collected in a human-facing study.

The idea is best illustrated with an example in Typed Racket's migratory type
system. Imagine a code base with dozens of modules in plain Racket. A developer
who opens a module for maintenance purposes must study the module's design and,
as part of the process, is bound to re-construct the types that went into the
module's creation.  To help future maintainers, the developer should report
these insights as type annotations. Over time, the code base migrates into a mix
of typed and untyped modules. As~\citet{tfffgksst-snapl-2017} report though,
it is equally common that developers add typed modules that depend on
the existing modules in the code base.

Now consider the concrete (and simplistic) example of figure~\ref{fig:rational}.
Initially the code base consists of the two \dyncolor\ modules on the left plus
the \typecolor\ module at the bottom; \dyncolor\ indicates untyped, while
\typecolor\ means typed. When a typed module imports an untyped module, it must
assign types to the imported identifiers for the type checker's sake. Here {\tt
main} specifies that {\tt na-client} consumes a {\tt Real} number and produces a
non-positive one.\footnote{Racket's type system reifies
reasoning about subsets of numbers, not machine-level
representations~\citep{stathff-padl-12}.} A program execution ends in this error:
\begin{verbatim}
    na-client: broke its own contract
      promised: (<=/c 0)
      produced: 40
      in: (-> any/c (<=/c 0))           
      contract from: (interface for na-client)
      blaming: (interface for na-client)
       (assuming the contract is correct)
\end{verbatim}
The referenced contract is the compilation of the type of {\tt na-client}. The
definitive hint is ``{\tt blaming: (interface for na-client)}'' with the caveat
``{\tt (assuming the contract is correct)}.''

Assuming the rational programmer trusts the type of {\tt na-client}, the next
step is to inspect the {\tt layer} module and to equip it with type
annotations. The result is the \typecolor\ module in the middle, and {\tt
main}'s import is now re-directed there by {\tt (submod ".." layer-typed)}. As
predicted by the theory, running the modified program (in the same way as before) yields a different error message:
\begin{verbatim}
    neg-abs: broke its own contract
      promised: (<=/c 0)
      produced: 10
      in: (-> any/c (<=/c 0))
      contract from: (interface for neg-abs)
      blaming: (interface for neg-abs)
       (assuming the contract is correct)
\end{verbatim}
Lastly, the rational programmer assigns types to {\tt server} and to re-direct
the import of {\tt layer-typed} to {\tt (submod ".." server-typed)}. Now the
type checker objects to the conjectured type of {\tt neg-abs}, i.e.\ the source
of the impedance mismatch is found. How to fix it is a separate question.

Like {\it homo economicus\/}, the rational programmer is an approximation.
People do not behave in a purely rational manner as economic actors, and they
also do not do so as software developers. The point is not to deny the existence
of ``lucky hunches'' programmers or ``tinkering works'' approaches and so on. It
is also not to claim that equipping entire modules with types represents an
always feasible approach\footnote{Adding types at the expression level, say, as
in TypeScript should be considered well within bounds.}  {\em But\/}, the
concept of studying the idea of an economically rational actor has produced
benefits to the discipline of political economics, and this paper suggests that
implementing and studying the rational programmer will help language designers.

%% -----------------------------------------------------------------------------
\subsection{How to Turn the Idea of the Rational Programmer into a Methodology} 
\label{sub:methodology}

\input{why-rational-methodology.tex}