### 2021-05-19 writing/paper.tex

\documentclass[acmsmall]{acmart}
% \settopmatter{printfolios=true,printccs=false,printacmref=false}

\citestyle{acmauthoryear}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from PACMPL format to traditional
%% SIGPLAN proceedings format must update the '\documentclass' and
%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Title information
\newcommand{\thetitle}{How to Evaluate Blame for Gradual Types}
\title{\thetitle}

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

\author{Lukas Lazarek}
\affiliation{%
  \institution{PLT @ Northwestern University}
  \city{Evanston}
  \state{Illinois}
  \country{USA}
}
\email{lukas.lazarek@eecs.northwestern.edu}

\author{Ben Greenman}
\orcid{0000-0001-7078-9287}
\affiliation{%
  \institution{PLT @ Northeastern University}
  \city{Boston}
  \state{Massachusetts}
  \country{USA}
}
\email{benjaminlgreenman@gmail.com}

\author{Matthias Felleisen}
\orcid{0000-0001-6678-1004}
\affiliation{%
  \institution{PLT @ Northeastern University}
  \city{Boston}
  \state{Massachusetts}
  \country{USA}
}
\email{matthias@ccs.neu.edu}

\author{Christos Dimoulas}
\orcid{0000-0002-9338-7034}
\affiliation{%
  \institution{PLT @ Northwestern University}
  \city{Evanston}
  \state{Illinois}
  \country{USA}
}
\email{chrdimo@northwestern.edu}
\renewcommand{\shortauthors}{Lazarek, Greenmanm Felleisen, Dimoulas}

\usepackage{alltt}
\usepackage{tcolorbox}
\usepackage{wrapfig}

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\usepackage{listings}
\usepackage[shortcuts]{extdash}
\usepackage{xcolor}

\usepackage{upgreek}
\usepackage{multirow}

\input{def.tex}

\acmConference[]{}{}{}
\acmYear{}
\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
%% \setcopyright{none}

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}

\begin{document}

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
Programming language theoreticians develop blame assignment systems and
  prove blame theorems for gradually typed programming languages.
  Practical implementations of gradual typing almost completely ignore the
  idea of blame assignment.  This contrast raises the question whether
  blame provides any value to the working programmer and poses
  the challenge of how to evaluate the effectiveness of blame assignment
  strategies. This paper contributes (1) the first evaluation method for
  blame assignment strategies and (2) the results from applying it to
  three different semantics for gradual typing. These results cast doubt to
  the theoretical effectiveness of blame. In most scenarios,  strategies
  with imprecise blame assignment are as helpful to a rationally acting
  programmer as strategies with provably correct blame.
\end{abstract}

% CCS concepts: Program specifications < Program reasoning < Semantics and Reasoning < Theory of Computation
% CCS concepts: Empirical software validation < Software verification and validation < Software creation and management < Software and its engineering

% potential choices decided against: pre- and post-conditions, assertions, software testing and debugging

%\begin{CCSXML}
%<ccs2012>
%<concept>
%<concept_id></concept_id>
%<concept_desc></concept_desc>
%<concept_significance></concept_significance>
%</concept>
%<concept>
%<concept_id></concept_id>
%<concept_desc></concept_desc>
%<concept_significance></concept_significance>
%</concept>
%</ccs2012>
%\end{CCSXML}

%\ccsdesc[]{}
%\ccsdesc[]{}

%% Keywords
%% comma separated list
%\keywords{}  %% \keywords are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

%% -----------------------------------------------------------------------------
\def\sec#1#2{\section{#2} \label{sec:#1} \input{#1.tex}}

\sec{introduction}{Does Blame Matter}
\input{landscape-program}
\sec{landscape}   {How to Think About Three Blame Systems}
\sec{why-rational}{What Is a Rational Programmer}
\sec{challenges}  {Why It Is Hard to Evaluate Blame}
\sec{rational}    {How to Make Comparable Rational Programmers}
\sec{mutate}      {How to Make Lots of Mistakes}
\sec{sample}      {How to Make the Experimental Space Tractable}
\sec{results}     {What Are the Outcomes of the Experiment}
\sec{discussion}  {What Can Programmers Learn}
\sec{related}     {Who has Worked on this Problem Before}
\sec{conclusion}  {What to Do Next}
%% -----------------------------------------------------------------------------

%% Acknowledgments
%%\begin{acks}                 
  %% acks environment is optional
  %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
%%\end{acks}

%% Bibliography
\bibliography{cs}

\end{document}