### 2021-05-20 writing/mutate-benchmarks.tex
move Louisiana footnote so that the intro to the benchmark section makes sense

%% -----------------------------------------------------------------------------

The benchmark programs for a rational-programmer experiment must 
\begin{enumerate}

\item vary in size, complexity and purpose;

\item be fully typed so that the choice of types is fixed;

\item take advantage of the variety of typing features of a gradually
typed; and

\item have a decent number of type-able modules and a variety of module dependency graphs
 because mixing of typed and untyped code in Typed Racket takes place at the
 module level.

\end{enumerate}

\citet{gtnffvf-jfp-2019}'s collection of Typed Racket programs for
systematically measuring the implementation's performance satisfies these
criteria. The benchmark suite consists of fully typed, correct programs, written
by a number of different authors, who have maintained and evolved these programs
over several years. The programs range widely in size, complexity, purpose,
origin and in programming style. The latter is mostly a reference to their
reliance on many Typed Racket features: occurrence typing~\citep{tf-icfp-2010},
types for mutable and immutable data structures~\citep{hpst-sfp-2010}, types for
first-class classes and objects~\citep{tsdtf-oopsla-2012}, and types for
Racket's numeric tower~\citep{stathff-padl-12}.  Finally, all of the programs
are deterministic, so any changes in the programs' behavior between runs can be
solely attributed to the actions of the rational programmer.

Hence, even if they do not come with type-level errors, they offer a diverse
starter suite, suitable for the careful introduction of type-level errors via
mutations. Further filtering the benchmark suite to keep the benchmarks with the
most sizeable and complex dependency graphs yields ten suitable programs while
preserving the suite's linguistic diversity. Figure~\ref{table:benchmark-descriptions} displays the selected
benchmark programs and describes their basic characteristics.