### 2021-05-20 writing/sample.tex

%% -----------------------------------------------------------------------------

%  from section~\ref{sec:mutate} (96,515,744)

As is, the chosen mutators generate approximately one million debugging
scenarios for the chosen benchmarks. This number of scenarios is far too large
to even identify the interesting ones among them.  Furthermore this population
is heterogeneous; scenarios come from different mutants, and the mutants are the
result of different mutators applied to each benchmark.

To render the experiment computationally feasible and statistically sound, it
becomes necessary to sample this large space in a uniform and stratified manner.
The first two levels of stratification group mutants first by benchmark and then
by mutator, making sure that the sample within each benchmark reflects the
diversity of mutants with respect to the mutators that generated them.
Specifically, the experiment samples 80 interesting mutants per benchmark,
evenly-distributed across all of the mutators that contribute mutants for the
benchmark. Some benchmarks have less than 80 mutants with interesting scenarios,
in which case the only choice is to include them all.  The result is a total of
752 interesting mutants across all benchmarks.  Finally, the third level of
sampling randomly draws 96 debugging scenarios from each configuration lattice
\mf{is ``with replacement'' correct here? Do you put it back for the next draw?}
with replacement. The final sample thus consists of 72,192 interesting
scenarios.