### 2021-06-01 writing/mutate.tex

%% -----------------------------------------------------------------------------

Putting the rational programmer to work means generating many mutants and turning
those into debugging scenarios. The process must start with a suitable
collection of representative programs (sec.~\ref{sub:mutate-benchmarks}).
Since existing mutators do not generate useful mutants, the next step is to
develop new mutators (sec.~\ref{sub:mutate-mutators}) and to validate their
suitability on the benchmarks (sec.~\ref{sub:mutate-interesting}).

%% -----------------------------------------------------------------------------
\def\sub#1#2{\subsection{#2} \label{sub:mutate-#1} \input{mutate-#1.tex}}

\input{mutate-benchmarks-figure.tex}
\sub{benchmarks}{The Experimental Benchmarks}
\input{mutate-mutators-figure.tex}
\sub{mutators}   {How to Mutate Software} 
\sub{interesting}{Are These Mutators Interesting}